\newpage
\section*{Abstract}

This thesis presents the design, implementation, and evaluation of an autonomous driving agent for the game \Trackmania{}. It builds on an earlier bachelor thesis prototype which showed that a Trackmania agent can perceive the track through exported geometry instead of raw screen images. The goal of this thesis is to turn that idea into a more systematic solution: to construct an environment representation, design the agent observation, automate policy training, and empirically evaluate the resulting driving behavior.

The proposed system reads the car state from the running game, projects it into a reconstructed virtual map, and computes a compact observation vector from the track geometry. The driving policy is represented by a multilayer neural network, and its parameters are searched primarily through neuroevolution with a genetic algorithm. Human driving data are first used to select a practical network architecture and to demonstrate the limitations of behavioral cloning. The thesis then compares several ways of evaluating driving behavior, including a lexicographic ranking function, reinforcement learning, and Pareto-based multi-objective comparison.

Within the performed experiments, the most practical main direction was lexicographic evaluation with the tuple \texttt{(finish, progress, -time, -crashes)} combined with genetic training. The resulting agent can finish tracks in a flat setting, on a track with height changes, and on a track with different surfaces. On the previously unseen map \texttt{small\_map}, the agent achieved a time of \(19.68\,\mathrm{s}\), while the bachelor-thesis agent achieved \(23.064\,\mathrm{s}\) and the reference human run achieved \(17.89\,\mathrm{s}\). In the tested conditions, these results support the claim that the new system clearly improves over the original prototype and that the geometric approach is useful beyond a simple flat track.

\paragraph*{Keywords:} Trackmania, autonomous agent, neuroevolution, genetic algorithm, neural network, virtual sensors
